Carbon Nucleosynthesis

### 2023-04-09 thesis/tex/main.tex
thesis corrections, to equilibrium

\usepackage{setspace}
\usepackage{array}
\setlength\extrarowheight{5pt}

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
% \patchcmd{<cmd>}{<search>}{<replace>}{<succes>}{<failure>}
\patchcmd{\@chapter}{\addtocontents{lof}{\protect\addvspace{10\p@}}}{}{}{}% LoF
\patchcmd{\@chapter}{\addtocontents{lot}{\protect\addvspace{10\p@}}}{}{}{}% LoT
\makeatother

@@

\usepackage{graphicx}	% Including figure files
\graphicspath{{../figures/}} 
\usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}
% \usepackage[titles]{tocloft} % fig titles
% \renewcommand{\cftfigpresnum}{\figurename\enspace}
% \setlength{\cftfignumwidth}{5em}
% \renewcommand{\cfttabpresnum}{\tablename\enspace}
% \setlength{\cfttabnumwidth}{5em}

@@

\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
\newcommand{\possesivecite}[1]{\citeauthor{#1}'s \citeyearpar{#1}}

@@

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
Stellar evolution models provide highly uncertain predictions for elemental yields due to our limited understanding of stellar physics. In this paper, I aim to estimate the nucleosynthetic yields of carbon using galactic chemical evolution models.

I find that the AGB stars make up a fraction $0.20_{-20}^{+80}$ of total carbon production at late times.

@@

Galactic chemical evolution (GCE) aims to understand the chemical and star formation history of galaxies. At the core of GCE models are stellar yields---the amount of each chemical element stars produce (or destroy). Different chemical elements are produced in different sites, so stars leave traces of the Galaxy's chemical history. Through \textit{Galactic Archeology}, we can reconstruct the history of our galaxy by investigating the clues left in the remains---stars from the early universe \citep{Weinberg+22, emily+19}.

I aim to understand the GCE of carbon, where the element is produced and how the abundances of this element evolve over time. Why stuy carbon? Carbon is unique nucleosynthetically, one of few light elements (along with nitrogen) which are produced in Asymptotic Giant Branch (AGB) stars \citep[for example,][]{jennifer19, KL14}. Carbon and nitrogen are well studied elements as they are easy to observe in stellar spectra, even at the lowest metallicities \citep{FN15} SOMETHING ELSE?. Additionally, carbon and nitrogen are used as age indicators in red giant branch stars \citep[e.g.][]{martig16, MG15, fiorenzo+21}.
While we know carbon is produced in AGB stars and core collapse supernovae (CCSNe), we still have limited understanding of the importaince and metallicity-dependence of each process.

We also know that there is evidence for non-monatonic carbon yield patterns. 
pattern where carbon is more readily produced by both high metallicity and very low metallicity stars (CEMPS, early GCE results), but our picture. \citep{FN15,cooke+17,berg+19}.

Stellar evolution models predict stellary yields, playing a crucial role in GCE.  However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, rotational mixing, opacity, and mass loss CITATIONS. Changes in opacity, rotation, reaction rates, and explodability can substantially affect carbon yields \citep{KL14,V13, LC18, emily+21}.

@@

While there are many stellar measurments of carbon, accurately determining the birth abundance is more challenging. For example, when a star becomes a RGB star, material from the CNO-processed core \citep{fiorenzo+21,KL14}. Since this core-processed material is now carbon depleted and nitrogen enhanced, measurments of this star's atmosphere will no longer reflect the birth abundances of the star.  Additionaly, carbon is faint in HII regions and determining gas-phase abundances is challenging \citet{skillman+20}.
Other surveys, such as GALAH, have unreliable carbon measurments with low detection rates \citep{emily+19}.

@@

In Fig. \ref{fig:subgiants}, I present a sample of APOGEE subgiants. These subgiants were chosen based on their location in the HR diagram based on the criteria in \citet{jack_subgiant} also replicated in Appendix \ref{sec:jack}. Because subgiants have not undergone FDU, their surface abundances are an accurate reflection of the birth abundances (CITATION).

\begin{figure}[htp]
    \includegraphics{subgiants_mgh.pdf}
    \includegraphics{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for our sample of APOGEE subgiants. On the top, I plot high-$\alpha$ stars in blue and low-$\alpha$ stars in orange using the seperation defined in Equation \ref{eq:high_alpha}. On the bottom, I colorcode stars according to their [Mg/H] abundance.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

Modern theoretical nucleosynthesis provides the starting point for our models.  I focus on three primary nucleosynthetic pathways: AGB stars, CCSne, and SNeIa. Each process has unique timescales and element production which I can trace through the tools of GCE. While carbon is only known to be produced in AGB stars and CCSNe, comparing carbon to iron production (from SNeIa) adds a valuable constraint to our models and insight into the delay-time production of carbon.

After a group of stars are formed, CCSNe from massive stars are the first enrichment, providing light elements such as C, O, and Mg and heavier elements such as Fe and beyond. Importaintly, $\alpha$ elements like O and Mg are only formed in CCSNe (in significant quantities). Next, low mass stars begin to reach the end of their lives. The dying breaths of AGB stars are known to be importaint sources of C, N, and a smattering of heavier elements.  Finally, the white dwarfs, including formation and time to collide/over-eat, explode, releasing much of Fe and nearby elements.

I focus on O and Mg as $\alpha$ elements only formed in CCSNe, and use Fe to
trace delayed enrichment from SNeIa. C and N are both formed in AGB and CCSNe.

There are two $\alpha$-elements I care about: Mg and O. Stellar observations
of magnesium are more reliable but oxygen abundances are easier to measure in
HII regions. I assume that both Mg and O are produced entirely in CCSNe with
constant, metallicity-independent yields. Thus $\text{[Mg/H]} = \text{[O/H]} =
[\alpha/\text{H}]$ in our models. (Observational validation?)

For carbon, the dominant producer is from massive stars which eject carbon fused in their cores during CCSNe.

To calculate the net initial mass function (IMF)-weighted yields, I use the \citet{james+21} model (hereafter \JJ) with the Versatile Integrator for Chemical Evolution code (\VICE).
As stellar evolution models predict per-star yields for stars of each mass, we need to integrate over the IMF to determine the total yields in terms of star formation.

As the focus of this work is on carbon yields, I keep all other yields fixed. The yields I use are based on \citet{james+21, james+22}.

Also following \citet{james+21, james+22}, I take the SNeIa Delay Time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

\begin{table}
	\centering
    \caption[Fiducial Model]{Yields for the fiducial model}
	\label{tab:fiducial_mod}
	\begin{tabular}{l l l l}
		\hline
        Element & $\Ycc$ & $\tilde{y}^{\rm AGB}$ & $y^{\rm SNe Ia}$ \\
		\hline
        C & $0.0028 + 0.001(Z/Z_\odot)$ & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & $9\times10^{-4}(Z/Z_\odot)M$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An AGB star is a low mass ($\lesssim 8 M_{\sun}$) star during the final phases of evolution.  In an AGB star, two competing processes determine the outcome of carbon production.
Third Dredge Up (TDU) is needed to eject carbon from the star, enriching the surface with helium-burning-processed material from the core. But Hot Bottom Burning (HBB) at the base of the convection layer changes carbon into nitrogen. 
TDU occurs during thermal pulsations, where the convective layer of the star passes into the inner layers with He-burning products which mixes in C and O into the envelope. In contrast, HBB may happen if the base of the convective envelope reaches about 50MK, initiating the CNO cycle, converting most C12 into N14.

Carbon abundances are strongly affected by CNO cycle processing in stars. The CNO cycle is one of two series of nuclear reactions( $^{12}$C(p, $\gamma$)$^{13}$N($\beta^+ \nu_e$)$^{14}$N(p, $\gamma$)$^{15}$O($\beta^+\nu_e$)$^{15}$N(p,$\alpha$)$^{12}$C

Because the $^{14}$N(p, $\alpha$) proton capture is the slowest component of the CNO cycle ****, the CNO cycle, to first order, converts all $^{12}$C into $^{14}N$.

The differences in yields between AGB models depend on the mass loss rate, treatment of convection, and nuclear reaction rates.

In this work, I explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item C11: \citet{cristallo+11} and \citet{cristallo+15}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16} \citet{karakas+18}
\end{itemize}
Unless otherwise noted, C11 is the model I choose as our fiducial AGB yield.

I use IMF-Integrated yields calculated with \VICE~throughout this paper (see
section \ref{sec:vice}, \citet{james+21, james+22}). For an element X and star
with mass $M$, the stellar yield $y$ is defined as the net production of X relative 
to $M$, or
\begin{equation}
    y_{\rm X} = \frac{M_{\rm X,\ ejected} - Z_{0, X} M_{\rm ejected}}{M}   
\end{equation}
where I let $M_{\rm ejected}$ and $M_{\rm X, ejected}$ be the total ejected
stellar mass (through winds and supernovae ejecta) and the ejected mass of X,
and $Z_{0, X}$ was the birth mass fraction of element X.

More useful for GCE are Initial Mass Function (IMF) yields, where I sum up the
contributions of all stars which produce an element.
I define the IMF-integrated yield of X with: 
\begin{equation}
Y_{\rm X}^{\rm proc}(Z) = \int_{M_{\rm min}}^{M_{\rm max}} y_{\rm X}^{\rm proc}(M, Z)\ \Phi(M)  \ dM
\end{equation}
Where $\Phi(M) = \frac{dN}{dM}/ \int_{M_{\rm min}}^{M_{\rm max}} \frac{dN}{dM}\ dM$ is the normalized IMF, $M_{\rm min}$ and $M_{\rm max}$ are the minimum and maximum mass of stars, which I take to be $0.08 M_{\sun}$ and $100 M_{\sun}$ respectively; and $\frac{dN}{dM}$ is the initial mass function, the frequency of stars born between $M$ and $M+dM$.

Fig. \ref{fig:y_agb} compares the net fractional AGB carbon yield, $m_\text{C}^\text{AGB}(M, Z)$ for each AGB model I consider. Note that the yields may be negative when the birth abundance is higher than the carbon abundance of the material ejected back to the ISM.

Most models agree on the qualitative shape of the net fractional AGB carbon yield. The stars with the highest fractional carbon yields are between 2 to 4 solar masses, depending on the metallicity and model. Both lower and higher mass stars produce less carbon, and high mass stars may even destroy carbon at high metallicity due to highly efficient HBB. And all models agree that the carbon yield generally decreases with metallicity.

Metallicity dependence:
V13 is the only model that stands out, showing a non-monotonic metallicity dependence, however this effect is only for models with [M/H]$\leq -1$. Otherwise the major differences between the models are the precise amount of carbon and the slope of the metallicity dependence. For example, the three models C11, K10, and K16 predict $y_\text{C}^\text{AGB}$ to be between 0.006 and 0.008 at solar metallicity, but C11 has a much shallower metallicity dependence than the K10 and K16 models. And at solar, V13 predicts a value at around 0.004 so the models are all within about a factor of 2.

Another difference is the speed at which equilibrium is reached. K10 and K16 weight carbon production more heavily towards massive stars resulting in a faster time to equilibrium, whereas the C11 and V13 models predict a slightly longer timescale with order 1Gyr, but little to no carbon is produced more than 2Gyr after a star formation event. This is in contrast to iron which is still produced by SNeIa up to 10Gyr after a star formation event.

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}\\

        \caption[AGB carbon yields]{The net fractional AGB carbon yield ($\tilde{y}$) plotted as a function of initial stellar mass $M$. Metallicity is colorcoded according to the colorbar. I plot $\tilde{y}=0$ for reference. Each panel represents yields from one of four AGB models: C11 \citep{cristallo+11, cristallo+15}, K10 \citep{karakas10}, V13 \citep{ventura+13,ventura+14,ventura+18}, and K16 \citep{KL16, karakas+18}.
            %C11 reports values at metallicities $Z=$ 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02; K10 at $Z=$0.0001, 0.004, 0.008, 0.02; V13 at $Z=$0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04; and K16 at $Z=$0.0028, 0.007, 0.014, 0.03.
        }
            \label{fig:y_agb}
\end{figure*}

\begin{figure}
    \includegraphics[scale=1]{y_agb_t2.pdf}

    \caption[AGB yields delay time distribution]{
        Carbon production by AGB stars as a function of SSP age, normalized to the total mass $M_{\rm tot}$ produced at $t=10$ Gyr. Left: The four AGB yield models from liturature  at solar metallicity (C11, K10, V13, or K16). The delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) is plotted as a dashed black line for comparison. Each model has a similarly shaped DTD, and most AGB enrichment occurs before 1 Gyr. Right: The C11 AGB model at different metallicities. With increasing metallicity, carbon enrichment occurs earlier, and the DTD eventually becomes non-monotonic.
    }

\begin{figure*}
    \centering

    \includegraphics[scale=1]{y_agb_vs_z.pdf}

    \caption[AGB yield metallicity dependence]{The net fractional IMF-weighted AGB carbon yield $\Ycagb$ as a function of metallicity for each of our AGB yield models.
    }

\end{figure*}

@@

Massive stars form $^{12}$C through the triple--$\alpha$ process. However, the carbon has to get out of the star to be useful. 
While there are many stellar models providing predictions of CCSNe yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

I summarise many of the predictions of stellar models in Fig. \ref{fig:y_cc}. 
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/O]^{CC}} = \log_{10}\left( \frac{y_{\rm C}^{\rm CC}}{y_{\rm O}^{\rm CC}}\right) - \log_{10} \left( \frac{Z_{C, \sun }}{Z_{O, \sun }} \right)
\end{equation}
Notice how even these ratios

    \begin{itemize}
        \item Wide range in predictions, exacerbated to 1 dex range when including variance in oxygen predictions
        \item (explain) most models have a near-metallicity independent oxygen yield, so I choose the fiducial value of 0.015 despite varience in $y_O^{cc}$
        \item The fiducial value, where I choose the carbon yield from Sukhbold+16, may be above most predictions of carbon yields but is within the range of yield ratios
        \item the fiducial value would reach an equilibrium carbon yield of -0.1 without AGB stars
        \item Rotation in LC18 appears to increase metallicity dependence of carbon production (read this paper more carefully)
        \item NKT and LC agree in positive metallicity depenent yields
    \end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics{y_c_cc.pdf}
    \caption[CCSNe carbon yields]{
        Carbon yields from massive stars.
        Top: The net fractional IMF-weighted CCSNe yield of carbon (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function) as a function of metallicity.
        Bottom: The CCSNe carbon to magnisium abundance ratio, defined in Equasion \ref{eq:c_mg_cc}. The black line represents the carbon yield of the fiducial model,
    $\Ycc = 0.0028 + 0.001 (Z/Z_{\odot})$. 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscales, while \citet{LC18} provide yields at different roational velocities.
The pink line represents the C11 AGB model.
}
    \label{fig:y_cc}
\end{figure}

@@

Chemical abundances reach an equilibrium. At least with simple enough assumptions. The new elements synthesized in stars balance against elements lost to galactic winds and locked in stellar reminants. While our galaxy is likely not in perfect equilibrium or described by a single, homogenous chemical region, I can still use an equilibrium approximation to understand total yields \citep[e.g.][]{james_dwarf,james+22,WAF17}.

>>>>

\usepackage{setspace}
\usepackage{booktabs}

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
% \patchcmd{<cmd>}{<search>}{<replace>}{<succes>}{<failure>}
\patchcmd{\@chapter}{\addtocontents{lof}{\protect\addvspace{10\p@}}}{}{}{}% LoF
\patchcmd{\@chapter}{\addtocontents{lot}{\protect\addvspace{10\p@}}}{}{}{}% LoT
\makeatother
\usepackage[tableposition=above]{caption}

@@

\usepackage{graphicx}	% Including figure files
\graphicspath{{../figures/}} 
\usepackage{natbib}
\usepackage[nottoc,numbib]{tocbibind}

\usepackage{isotope}
% \usepackage[titles]{tocloft} % fig titles
% \renewcommand{\cftfigpresnum}{\figurename\enspace}
% \setlength{\cftfignumwidth}{5em}
% \renewcommand{\cfttabpresnum}{\tablename\enspace}
% \setlength{\cfttabnumwidth}{5em}

@@

\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
\newcommand{\possesivecite}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand{\alp}{$\alpha$}

@@

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
Stellar evolution models provide highly uncertain predictions for elemental yields due to our limited understanding of stellar physics. In this paper, I aim to estimate the nucleosynthetic yields of C using galactic chemical evolution models.

I find that the AGB stars make up a fraction $0.20_{-20}^{+80}$ of total C production at late times.

@@

Galactic chemical evolution (GCE) aims to understand the chemical and star formation histories of galaxies. At the core of GCE models are stellar yields---the amount of each chemical element stars produce (or destroy). The chemical elements are produced in different amounts by different stars, leaving Galaxy's evolutionary history. Through \textit{Galactic Archeology}, we can reconstruct the history of our galaxy by investigating the clues left in reminants from the past \citep{Weinberg+22, emily+19}.

Here, I aim to understand C enrichment---where it is produced and how its abundances evolve over time. Why study C? C is unique nucleosynthetically, one of few light elements (along with N) produced in Asymptotic Giant Branch (AGB) stars \citep[e.g.][]{jennifer19, KL14}. C and N are well studied elements as they are easy to observe in stellar spectra, even at the lowest metallicities \citep{FN15} SOMETHING ELSE?. Additionally, C and N are used as age indicators in red giant branch stars \citep[e.g.][]{martig16, MG15, fiorenzo+21}.
While we know C is produced in AGB stars and core collapse supernovae (CCSNe), we still have limited understanding of the importaince and metallicity-dependence of each process.

Empirically, gas-phase [C/O] ratios decrease with increasing [O/H] at the lowest metallicitis---in high redshift damped Lyman-alpha systems (DLAs)---and then [C/O] increases above $\rm [O/H] \approx 1$ (\citealt{berg+19,cooke+17}; see discussion in SECTION).

\cite{james+22} examend GCE models of N, an elements whose production is closely related to C. \cite{james+22} found trends in N and O abundances can be explained by the metallicity dependence of relatice N and O yields. Here, I extend their models to C and show that similar constraints on C and O relative yields can be obtained from observed abundance ratios.

We also know that there is evidence for non-monatonic C yield patterns. 
pattern where C is more readily produced by both high metallicity and very low metallicity stars (CEMPS, early GCE results), but our picture. \citep{FN15,cooke+17,berg+19}.

Despite their centrol role in GCE models, nucleosynthetic yields predicted by stellar evolution models are rife with uncertainties. Poorly understood processes, including mass loss, opacity, nuclear reaction rates, rotational mixing, and convection/convective boundaries each shape the production rates of atomic nuclei in stellar interios \citep{KL14,V13, LC18, emily+21}.
Additionally, changes in massive star's explodabinitily can introduce a factor of \~2 change in population averaged-yields \citep{emily+21}.

Stellar evolution models predict stellary yields, playing a crucial role in GCE.  However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, rotational mixing, opacity, and mass loss CITATIONS. Changes in opacity, rotation, reaction rates, and explodability can substantially affect C yields

@@

While there are many stellar measurments of C, accurately determining the birth abundance is more challenging. For example, when a star becomes a RGB star, material from the CNO-processed core is mixed with the envelope in First Dredge Up (FDU) \citep{fiorenzo+21,KL14}. Since this core-processed material is now C depleted and N enhanced, measurments of this star's atmosphere will no longer reflect the birth abundances of the star.  Additionaly, C is faint in HII regions and determining gas-phase abundances is challenging \citet{skillman+20}.
Other surveys, such as GALAH, have unreliable C measurments with low detection rates \citep{emily+19}.

@@

To this end, I use a sample of subgiant stars from APOGEE \citep{apogee17}. According to stellar evolution models (CITES), these stars have not yet experienced dredge-up. Their photospheric abundances should thus represent their birth composition.  In Fig. \ref{fig:subgiants}, I plot this sample, selected by criteria in \citet[see Appendix \ref{sec:jack}]{jack_subgiant}. These emperical resolults contain information on the relative C and O yields, which I use later in this work.

\begin{figure}[htp]
    \includegraphics{subgiants_mgh.pdf}
    \includegraphics{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for the \citet{jack_subgiant} sample of APOGEE subgiants. On the top, I plot high-$\alpha$ stars in blue and low-$\alpha$ stars in orange using the seperation defined in Equation \ref{eq:high_alpha}. On the bottom, I colorcode stars according to their [Mg/H] abundance.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

Theoretical models of stellar nucleosynthesis provide the starting point of this investigation. I focus on three primary nucleosynthetic pathways: Asymptotic Giant Branch (AGB) stars, Core Collapse Supernovae (CCSNe), and Type Ia Supernovae (SNe Ia). Each process has unique timescales and yields, tracable through the tools of GCE. While C is only produced in AGB stars and CCSNe, comparing C to Fe, which is produced in SNe Ia, adds a valuable constraint on the delay-time distribution (DTD) of C enrichment.

When a Single Stellar Population (SSP) forms, CCSNe are the first enrichment events. CCSNe provide light elements such as C, O, and Mg and heavier elements such as Fe and beyond. Metallicity-independent yields from CCSNe are the only statisticly significant source of O and Mg ($\alpha$ elements). Next, low mass stars begin to reach the end of their lives. The dying breaths of AGB stars are known to be importaint sources of C, N, and neutron capture elements.  Finally, white dwarfs explode, releasing Fe and other iron-peak elements.

Stellar measurments of Mg are more reliable, but O abundances are easier to measure in HII regions. 
I therefore use Mg abundances from the \citet{jack_subgiant} subgiant sample as my primary emperical constraint. However, I use O abundances for gas-sphase measurments and low-metallicity stars (\S \ref{sec:gas}). I model both Mg and O as metallicity-independent population-averaged yields dominated by massive stars. ${\rm [O/Mg]}\approx 0$ across a wide range of metallicities in APOGEE \citep{Weinberg+19, Weinberg+22}. 
I adopt O and Mg relative yields given by the solar mixture of \cite{asplund+09} <MATH>.

To calculate the IMF-weighted\footnote{IMF: Initial Mass Function} net yields of C, I use the Versatile Integrator for Chemical Evolution code (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}).
As stellar evolution models predict per-star yields of a given mass, I need to integrate over the IMF to determine the total yields in terms of star formation.

In the interest of focusing on carbon yields, I hold yields of other elements fixed. 
I adopt yields from \citet{james+21, james+22}.
In Table \ref{tab:fiducial_mod}, I tabulate yield values, in units of a stellar population's birth mass.
Also following \citet{james+21, james+22}, I take the SNeIa Delay Time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

\begin{table}
	\centering
    \caption[Fiducial Model]{Yields for the fiducial model}
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l}
		\toprule
        Element & $\Ycc$ & $\tilde{y}^{\rm AGB}$ & $y^{\rm SNe Ia}$ \\
		\midrule
        C & $0.0028 + 0.001(Z/Z_\odot)$ & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & $9\times10^{-4}(Z/Z_\odot)M$ & 0\\
		\bottomrule
	\end{tabular}
\end{table}

@@

An AGB star is a low mass ($\lesssim 8 M_{\sun}$) star during its final phase of evolution.  In an AGB star, two competing processes determine the outcome of C production: Third Drege Up (TDU) and Hot Bottom Burning (HBB).  DU accompianies thermal puslations in AGB stars. Durring a TDU event, material from the CO core is mixed with material exposed to partial He-shell burning, increasing C and O abundances in the envelope \citep{KL14}.
HBB is the activation of proton capture reactions, and the CNO cycle%
\footnote{The CNO (I) cycle is
\isotope[12]{C}(p, $\gamma$)
\isotope[13]{N}($\beta^+ \nu_e$)
\isotope[13]{C}(p, $\gamma$)
\isotope[14]{N}(p, $\gamma$)
\isotope[15]{O}($\beta^{+}\nu_e$)
\isotope[15]{N}(p, \alp)
\isotope[12]{C}.
 Other minor branches of the CNO cycle are not relevant here.
 \citep{solar-fusion}.
}
at the base of the convective envelope. Because the $^{14}$N(p, $\alpha$) proton capture is the slowest component of the CNO cycle \citep{solar-fusion}, the CNO cycle converts nearly all \isotope[12]{C} into \isotope[14]{N}.
As a result, when both TDU and HBB occur, \isotope[12]{C} yields are lowered (see discussion in \citealt{james+22} and \citealt{ventura+13}).

In this work, I explore four different sets of AGB star yield tables from literature, providing well sampled grids in metallicity and mass for use in chemical evolution models. 
\begin{itemize}
    \item C11: \citet{cristallo+11} and \citet{cristallo+15}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16} \citet{karakas+18}
\end{itemize}
I use C11 as the fiducial AGB yield.
For an element X and star with mass $M$, the stellar yield $\tilde{y}$ is defined as the net production of X relative to $M$, or
\begin{equation}
    \tilde{y}_{\rm X} = \frac{M_{\rm X,\ ejected} - Z_{0, X} M_{\rm ejected}}{M}   
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm X, ejected}$  are the total ejected mass of the envelope and the element X, respectively.

Fig. \ref{fig:y_agb} compares the AGB C yield $\tilde{y}_\text{C}^\text{AGB}(M, Z)$ for each model I consider. 
The yields may be negative when the birth abundance is higher than the C abundance of the material ejected back to the ISM. 
Most models agree on the qualitative shape of the net fractional AGB C yield---% 
stars with mass $\sim 2--4 M_\odot$ have highest fractional C yields, with the mass of the peak increasing and overal yields decreasing with increasing $Z$.  High mass, high $Z$ stars destroy \isotope[12]{C} due to HBB.

In Fig. \ref{fig:agb-ssp}, we plot the total production of C by AGB stars as a function of SSP age.
Effectively, this is equivalent to integrating Eq. \ref{eq:imf-yield} down to $M_{to}$, the turnoff mass for time $t$. 
\footnote{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr,
from \citet{larson74}.}
As the mass range $2\lesssim M \lesssim 4$ is most important for C production, the characteristic delay time is $\sim 1$Gyr. In contrast, Fe from SNeIa (plotted as the dashed black line in Fig \ref{fig:agb-ssp}) has a longer characteristic delay time.

K10 and K16 weight C production more heavily towards massive stars resulting in a faster enrichment delay time, whereas the C11 and V13 models predict a slightly longer timescale with order 1Gyr, but little to no C is produced more than 2Gyr after a star formation event. This is in contrast to iron which is still produced by SNeIa up to 10Gyr after a star formation event. 
As shown in the right panel of Fig. \ref{fig:agb-ssp}, with increasing $Z$, C enrichment occurs earlier, and C destruction in low mass stars leads to a declining C abundance at late times.

\begin{figure}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[AGB C yields]{The net fractional AGB C yield ($\tilde{y}$) plotted as a function of initial stellar mass $M$. Metallicity is colorcoded according to the colorbar. I plot $\tilde{y}=0$ for reference. Each panel represents yields from one of four AGB models: C11 \citep{cristallo+11, cristallo+15}, K10 \citep{karakas10}, V13 \citep{ventura+13,ventura+14,ventura+18}, and K16 \citep{KL16, karakas+18}.
        }
        \label{fig:y_agb}
\end{figure}
\begin{figure}
    \includegraphics[scale=1]{y_agb_t2.pdf}

    \caption[AGB yields delay time distribution]{
        C production by AGB stars as a function of SSP age, normalized to the total mass $M_{\rm tot}$ produced at $t=10$ Gyr. Left: The four AGB yield models from literature at solar metallicity (C11, K10, V13, or K16). The delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) is plotted as a dashed black line for comparison. Right: The C11 AGB model at different metallicities. }
    \label{fig:agb-ssp}
\end{figure}

Though per-star yieldds are necessary to compute AGB star enrichment rates in GCE models, IMF-averaged yields are useful in interpreting their predictions. For a yield $y$ from a star of mass M and initial metallicity $Z$, the IMF-averaged yield is given by 
More useful for GCE are Initial Mass Function (IMF) yields, where I sum up the
contributions of all stars which produce an element.
I define the IMF-integrated yield of X with:

\begin{equation} \label{eq:imf-yield}
    y_{\rm X}^{\rm proc}(Z) = 
    \int_{M_{\rm min}}^{M_{\rm max}} 
    \tilde{y}_{\rm X}^{\rm proc}(M, Z)
    \frac{dN}{dM}  \ dM
\end{equation}

where $\frac{dN}{dM}$ is the normalized IMF, $M_{\rm min}$ and $M_{\rm max}$ are the minimum and maximum mass of stars, which I take to be $0.08 M_{\sun}$ and $100 M_{\sun}$ respectively. For AGB stars, the yields are truncated above 8 $M_{\odot}$.

In Fig. \ref{fig:yagb-z}, I plot the IMF-averaged C yields for each AGB model as a function of metallicity $Z$.
V13 differs that it shows a non-monotonic metallicity dependence. However, this effect is only form models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models only differ in the yield normalization and strength of metallicity dependence. All models predict yields within a factor of $\sim 2$.
For example, the three models C11, K10, and K16 predict $y_\text{C}^\text{AGB}$ to be between 0.006 and 0.008 at solar metallicity, but C11 has a much shallower metallicity dependence than the K10 and K16 models. V13 instead predicts a yield $\sim 0.004$.

\begin{figure}
    \centering
    \includegraphics[scale=1]{y_agb_vs_z.pdf}

    \caption[AGB yield metallicity dependence]{The net fractional IMF-weighted AGB C yield $\Ycagb$ as a function of metallicity for each of our AGB yield models.
    }
    \label{fig:yagb-z}
\end{figure}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process. However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of CCSNe yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

In Fig. \ref{fig:y_cc}, I plot calculations of the IMF-integrated yields, using Eq. \ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). On the right, I also show the CCSNe-[C/Mg] ratio, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
In other words, ${\rm [C/Mg]^{CC}}$ describes what [C/Mg] would be if CCSNe were the only process producing C.
Notice that even at $Z=Z_\odot$, there is a wide range ($\sim 1$ dex) in predicted $\Ycc$.

    \begin{itemize}
        \item Massive stars should dominate C yields, GCE models indicate similar results \citep[e.g.][]{prantzos+18,rybizki+17}.  
        \item Rotation in LC18 appears to increase metallicity dependence of C production (read this paper more carefully)
        \item NKT and LC (with rotation)agree in positive metallicity depenent yields
        \item [C/Mg]
        \item The fiducial value, where I choose the C yield from Sukhbold+16, may be above most predictions of C yields but is within the range of yield ratios
        \item the fiducial value would reach an equilibrium C yield of -0.1 without AGB stars
    \end{itemize}

\begin{figure}[htp]
    \centering
    \includegraphics{y_c_cc.pdf}
    \caption[CCSNe C yields]{
        C yields from massive stars.
        Top: The IMF-weighted CCSNe yield of C a function of metallicity.
        Bottom: The CCSNe [C/Mg] abundance ratio, defined in Eq. \ref{eq:c_mg_cc}. The black line represents the C yield of the fiducial model,
    $\Ycc = 0.0028 + 0.001 (Z/Z_{\odot})$. 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscales, while \citet{LC18} provide yields at different roational velocities.
The pink line represents the C11 AGB model.
}
    \label{fig:y_cc}
\end{figure}

@@

In the presence of metal-poor gas accretion and feedback-driven outflows, galaxies reach an equilibrium abundance in which production of new metals is balanced by losses to new stars and outflows \citep{larson72, dalcanton07, FD08, PS11, lilly13}.
While our galaxy is likely not in perfect equilibrium or described by a single, homogenous chemical region, the equilibrium approximation is nevertheless useful in understanding yields and metallicity dependence \citep[e.g.][]{james_dwarf,james+22,WAF17}.

I assume a simple ``one-zone'' chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  The gas and stars are homogenously and instantaneusly mixed, so spacial dependence is neglected.
